\PYGEOMETRY{}: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2021-12-09 public/paper/sub-cpc-2/reviewer_comments_response.txt

Authors : The authors would like to thank the reviewers for their comments. Many of the issues found improve the readability of the manuscript for potential users of the software. Almost all of the comments and suggested changes were implemented as suggested by the reviewers. Below are the reviewer comments with the changes or comments.

Reviewer #1: The authors are to be commended on writing an exceptionally clear and comprehensive manuscript. While the paper contains occasional grammatical errors, it is easy to follow and addressed all concerns I had as I was reading it.
This paper concerns pyg4ometry, a tool to create Geant4 geometry structures in GDML via Python, with tools that support restructuring files in CAD and FLUKA formats. Examples are provided throughout the manuscript to aid the reader in implementing the module,
and docs are already available online for v0.9.1 as of this review. As a former user of GATE, I am personally aware of the difficulty in rapid creation of Geant4 geometries, and I wish I had such a tool at my disposal in the past.
This work helps lower the barrier of entry to Geant4 for inexperienced programmers who do know enough Python to get by. I think a line expressing this is warranted.
    Line added

This manuscript is ready for publication as written, but I leave the following comments for the authors’ consideration:
P3 Line 27: powerful how?
   Removed word powerful   
P4 Line 9: I'd add a line highlighting how useful it is that this co
de is written for Python users. ROOT, for example, is harder for a layman to use.
   Added extra line to clarify this
P4 Line 56: stick a comma before which or replace it with that. (Come to think of it, I don't know if American grammar rules apply in the UK)
   Fixed
P7Line 50: I think this entire paragraph should be placed in a discussion section and not in methods. Best not to promise things for future release in a methods section.
   Changed as loops are unlikely to be implemented.
P10 L37: Too not to. In fact, this whole paragraph is oddly out of place so far. Use of clearly and loopy language that is less specific and efficient than its preceding chapters. What is it trying to say?
    Paragraph tightened up for clarity

Reviewer 2 [Linnemann]
The manuscript is generally well-written, though the English could use touching up: there are a lot of run-on sentences. I recommend that the manuscript be accepted but that the authors incorporate some minor revisions for the final manuscript.
Suggestions below (page.line)

2.16 impossible to convert -> impossible
    Fixed
2.24 STL undefined
    Fixed
2.33 delete: different
    Deleted
2.34 material, such -> material. Such
    Fixed
2.39 specialism -> specialty
    Fixed
2.59 geometry -> geometry descriptions
    Fixed
3.36 toolkit. A user
    Fixed
4.21 using a lot of -> using
    Fixed
4.29 existent -> existing
    Fixed     
4.31 follows: first comes
    Changed
4.46 delete: different
    Fixed
In this section it might make sense to mention that you chose STP and STEP as manufacturer-neutral CAD file formats for which there commonly exist converters from proprietary formats.
And that IGES is an older standard which STEP seems to be overtaking.
But it appears that Fig 1 is missing IGES since you later state IGES can be read, which is confusing since at 6.47 you omit IGES, but include it at 30.49; perhaps you mean that the software mentioned there can translate IGES to STEP for you?
I also note you don't mention OBJ why? Should it be mentioned? Any plans for adding in the future?
    OBJ is mentioned and other similar formats are trivally implemented, we provide one simple one. 
I of course wondered whether AUTOCAD is compatible, but apparently there are converters from DWG default to STEP, so no need to mention that specific product if you make the more general statements. But aha at 32.41 you mention another piece of Autodesk software, but state that you can save as STEP format.
    Autodesk FUSION replaced with Common commercial program. Wanted to make the point that the input for pyg4ometry was from a standard package a user might be familiar with 
5.10 solids. However,
    Fixed
8.43 geometry, a rapid
    Fixed
9.56 user. It does
    Fixed
10.37 to -> too
    Fixed
10.48 reuse. However
    Not changed. 
10.53 considered. In
    Fixed
13.13 vistualisation -> visualization (assuming British spelling)
    Fixed
15.18 trivial. A
    Fixed
18.33 50km is there a place this default can be changed? I know I'm weird, but I'm thinking astrophysics
    This can easily be changed and note added to highlight this.
Quadric hmm, maybe usage has changed since I studied mathematics I'd have said quadratic, but quadric seems correct when I look it up!
21.14 straight forwards -> straightforward
    Fixed
22.46 geometries. However
    Fixed
29.14 doing do -> doing so
    Fixed
30.9 PLA. It is
    Fixed
32.33 assemblies. These
    Fixed
32.41 I'd remove comma after electromagnet
Hmm, now you mention another piece of Autodesk software, but do state that you can save as STEP format
    Agreed and fixed
32.42 STEP file. The PYG4OMETRY
    Fixed    
32.40 PYG4OMETRY. This
    Fixed
33.52 FLUKA geometry, is converted
    Fixed
34.19 developer-written falls outside the margins
    Fix in proof (as likely to change)
34.52 FLUKA,
    Fixed     
35.24 user that -> users who
    Fixed    
35.39 hmm, apparently ParaView is a GUI for VTK; if you combine them into one object, I'd have said is instead of are
    Fixed

******************************************
Please note that the editorial process varies considerably from journal to journal. To view a sample editorial process, please click here:
http://ees.elsevier.com/eeshelp/sample_editorial_process.pdf
For further assistance, please visit our customer support site at http://help.elsevier.com/app/answers/list/p/7923. Here you can search for solutions on a range of topics, find answers to frequently asked questions and learn more about EM via interactive tutorials. You will also find our 24/7 support contact details should you need any further assistance from one of our customer support representatives.~

### 2021-12-09 public/paper/sub-cpc-2/pyg4ometry.tex

%\linenumbers

@@

\begin{small}
\noindent
{\em Program Title: \PYGEOMETRY{} }                                         		\\
{\em Licensing provisions: GPLv3 }							\\
{\em Programming language: Python, C++}                         		\\
{\em External routines/libraries: ANTLR, CGAL, FreeCAD, NumPy, OpenCascade, SymPy, VTK} \\
%{\em Supplementary material:}                                 				\\
  % Fill in if necessary, otherwise leave out.
%{\em Journal reference of previous version:}                  			\\
  %Only required for a New Version summary, otherwise leave out.
%{\em Does the new version supersede the previous version?:}   	\\
  %Only required for a New Version summary, otherwise leave out.
%{\em Reasons for the new version:}							\\
  %Only required for a New Version summary, otherwise leave out.
%{\em Summary of revisions:}*								\\
  %Only required for a New Version summary, otherwise leave out.
{\em Nature of problem:}\\  % (approx. 50-250 words)
Creating computer-readable geometry descriptions for Monte Carlo radiation transport (MCRT) codes is a time-consuming and error-prone task.
Typically these geometries are written by the user directly in the file format used by the MCRT code. There are also multiple MCRT codes
available and geometry conversion is difficult or impossible to convert between these simulation tools.
\\
{\em Solution method:}\\  % (approx. 50-250 words)
Create a Python application programming interface for the description and manipulation of Geant4 and FLUKA geometries, with full support for
the direct reading and writing of their respective geometry description file formats.
Form triangular meshes to represent geometric objects for both visualisation of the
geometry and to enable the use of advanced mesh-based geometric algorithms. Triangular mesh processing algorithms allow the loading and use of STL and CAD/CAM
files. Converting from FLUKA to Geant4 requires algorithms to decompose solids to a set of unions of convex solids. Converting from
FLUKA to Geant4 requires a number of steps including the replacement of infinite surfaces with finite solids and the automatic elimination of overlaps.

@@

%% main text
\section{Introduction} \label{sec:introduction}
There are numerous different software codes to simulate the passage of particles through material, such radiation transport (RT) programs
include MCNP~\cite{Mcnp_Werner}, FLUKA~\cite{Fluka_Ferrari,Fluka_Bohlen}, Geant3~\cite{Geant3_Brun} and Geant4~\cite{Geant4_Agostinelli}.
All these codes are based on the Monte Carlo technique but each code either has a particular specialism, simulation methodology or target user community.
 Monte Carlo RT (MCRT) simulations have diverse uses including shielding calculations for radiological protection, detector performance, medical
imaging and therapy, and space radiation environment simulations. A fundamental requirement of all of the codes is to supply a
computer-readable description of the  physical 3D geometry that  the particles are passing through.  The creation of geometry files is
typically a very time-consuming activity and the simulation validity and performance is directly dependent on the quality of the geometry. There is no
standard geometry format used across MCRT codes, with each code often using its own unique format. A user will typically not have geometry in both FLUKA and Geant4 for example. A geometry system
that enables the conversion between files prepared for different codes will allow for cross-checks of the physics processes in different particle transport
 codes.  The file formats used for geometry are generally focused
 on the computational efficiency of particle tracking algorithms  and not ease of preparation. In addition to the creation of geometry files for RT programs, usually
 computer-aided design (CAD) files exist for systems which need to be simulated. The fundamental geometric representations in CAD files are usually not
 amenable to MCRT programs.  For these reasons it is advantageous to create a software tool that allows particle transport code users to rapidly develop
 error-free geometry files, convert between common MCRT geometry formats and load CAD models.

This paper describes a geometry creation and conversion package called \PYGEOMETRY{}, written in Python and internally based on the Geant4 application
programming interface (API) and the Geometry Description Markup Language (GDML) for file persistency~\cite{GDML}. The main features of \PYGEOMETRY{}
are a Python scripting API to rapidly design parametrised geometry; conversion to and from  FLUKA geometry descriptions; conversion from CAD formats  (STEP
and IGES) based on FreeCAD~\cite{FreeCAD} and OpenCascade~\cite{OpenCASCADE}; and powerful geometry visualisation tools based on VTK~\cite{VTK4}. The
origin of \PYGEOMETRY{} was a set of utilities to prepare geometry for an accelerator beamline simulation program based on Geant4 called BDSIM~\cite{BDSIM_Nevay}.
Accelerator physicists, like specialists in other areas, need a tool to quickly model specialist geometry and the subsequent interaction of the charged particle beam.
\PYGEOMETRY{} allows the rapid creation and adaptation of geometry, with Figure~\ref{fig:workflow} demonstrating various possible workflows. \PYGEOMETRY{}
is not an executable software package but a toolkit, a user would typically write a very small Python program to use the classes and functions provided by
\PYGEOMETRY{}. This paper describes version 1.0 of \PYGEOMETRY{}, which is freely available as a Git repository and via the Python Package Index (PyPi).

@@

There are existing codes that have functionality similar to \PYGEOMETRY{}. ROOT~\cite{Brun:1997pa}, the high-energy physics data analysis framework,
can load, display and manipulate GDML-like geometry. Existing open source tools to convert CAD files to GDML include GUIMesh~\cite{GUIMesh_Pinto}, with commercial
solutions including ESABASE2~\cite{ESABASE2} and FASTRAD~\cite{FASTRAD}. There are also tools from the fusion and neutronics community that can convert CAD
geometry into formats usable by MCRT codes, with examples including DAGMC~\cite{DAGMC} and McCAD~\cite{McCad}. In principle CAD software can export shape data to STL
(or other similar mesh formats), which can be used by Geant4~\cite{poole2012acad}. However, in practice using a lot of CAD models is difficult if that model is composed of a
large number of parts. This is because exporting, assigning material to, and placing the STL components into the MCRT code can be very cumbersome. Almost all modern CAD tools such as
CATIA, Inventor and SolidWorks have a scripting language to allow users to programmatically generate geometry. Similar CAD-style scripting languages do not exist for either GDML
or FLUKA and the existent set of software does not provide a complete set of tools to efficiently create complex geometries for these codes.

This paper is structured as follows, first a brief radiation
transport-focused introduction to computer descriptions of geometry and an explanation of the design
and implementation of \PYGEOMETRY{}. Subsequent sections describe how \PYGEOMETRY{} can be used to perform rapid geometric modelling as well as conversions from
FLUKA to GDML, GDML to FLUKA and CAD to GDML. The paper concludes with an example of a composite, complex system consisting of components drawn from
all the supported  geometry input files formats.

\section{Computer descriptions of geometry} \label{sec:geometric}
Central to a computer-readable geometry is how a solid is defined in three dimensions. There are numerous different ways to describe a
geometry, including constructive solid geometry (CSG), boundary representation (BREP) and tessellated polygons, which are described
briefly in this section.  The Geant4 geometry specification is a mixture of all three of these geometry modelling techniques and described in detail last.

%%%%%%%%%%%%%%%%% Constructive solid geometry (CSG)
Constructive solid geometry uses Boolean operations (subtraction, intersection and union) between simple solid shapes (e.g. cube, cylinder, sphere, etc.) or infinite
volumes (e.g. a plane-defined infinite half-space) to model complex surfaces which represent a solid. Boolean operations and solids can be combined to form a
CSG tree to model complex geometry. FLUKA uses CSG to model solids, however the form of Boolean expression used by FLUKA is not a general CSG tree but a
logical expression in disjunctive normal form.

@@

\subsection{Visualisation} \label{sec:visualisation}
When implementing geometry a rapid and robust visualisation system is key to produce error-free and efficient simulation input.
A  \PYGEOMETRY{} geometry hierarchy can be viewed using the popular Visualisation Toolkit (VTK). No separate scene graph is required as the Geant4
volume hierarchy is sufficient to place the meshes associated with each physical volume. A daughter volume is placed within a logical volume with a
rotation~$\mathbf{R}_d$, reflection~$\mathbf{S}_d$ and translation~$\mathbf{T}_d$.

@@

\subsection{Overlap detection}
All MCRT codes cannot handle spatial overlap between two geometric objects and will have ill-defined behaviour when tracking particles
in such a situation.  A key feature of \PYGEOMETRY{} is the detection of potential overlaps in a way which is most useful to the user, it does this by
performing an intersection operation between solid instances and determining if the resulting mesh is empty. Figure~\ref{fig:overlap} shows three different types of possible
overlaps, (a) protrusion of a daughter from the mother, (b)  finite volume intersection between two daughters  and (c) an overlap where two daughters
share a face. If the resulting intersection is non-null then the overlaps can be displayed side-by-side in the visualisation.
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=8cm]{overlap}
\caption{Schematic of the three different types of overlap between the daughters of a mother logical volume.}
\label{fig:overlap}
\end{center}
\end{figure}

@@

Overlap detection in \PYGEOMETRY{} relies on the meshes generated for each solid.  As
these mesh representations will generally approximate their respective solid,
so to will any overlap detection algorithm be an approximation. Generally, the overlap detection
algorithm proceeds as shown in Algorithm~\ref{algo:overlap}.
For a logical volume with $n_{\rm daughters}$ physical volumes, assuming meshes for solids have an average number of faces $n$, clearly this algorithm
has complexity $\sim \mathcal{O}(n^2 n_{\rm daughters}^2)$. This complexity
can result in a steep computational cost for highly-flat geometry
hierarchies, but overall remains worthwhile considering the potential
waste in large amounts of cluster CPU time if small overlaps are present in the final MCRT simulation. This algorithm clearly favours geometry
descriptions which have a high degree of logical volume reuse, however this is also true of Geant4 as a whole, so the user will likely be inclined to design along such lines regardless.
Due to the discrete nature of triangular meshes it is not possible to have perfect detection
of overlaps, especially when curved surfaces are considered, and in some rare
cases either overlaps maybe missed, or spurious overlaps maybe reported.  However, the density of the meshes
created for the solids is controllabe by the user, meaning that the user
can opt for more precise overlap checking at the cost of greater
computation time.  In general, the overlap detection algorithm can
present the potential overlaps quickly and easily to the user, thus
significantly aiding the design process.

@@

# physical volume vistualisation attributes
bp1.visOptions.color = (1,0,0)
bp1.visOptions.alpha = 1.0
bp2.visOptions.color = (0,1,0)
bp2.visOptions.alpha = 1.0
bp3.visOptions.color = (0,0,1)
bp3.visOptions.alpha = 1.0

@@

The \PYGEOMETRY{} Python code in the example is approximately as expressive
as the GDML it writes. The benefit of wrapping GDML in Python is that it allows
 very rapid prototyping of geometry without the overhead of C++ compilation  (in the
case of implementing the geometry directly in Geant4) or writing well-formed XML (in the case of GDML). Effectively,
by using \PYGEOMETRY{}, the set of possible user-errors when describing geometry in C++ or XML
instead either manifest as Python exceptions or are eliminated entirely.
Another key benefit is the ability to use the \PYGEOMETRY{} code to create
programmatic converters between different geometry languages or more generally
manipulation and transformation of the geometry stored in memory. The rapid modelling
example given in Listing~\ref{lst:pythonRapidModelling} and Figure~\ref{fig:rapidModellingExample}
is rather trivial, a significantly more complex example is shown in Figure~\ref{fig:gdml-flair}.

@@

Generating these bounding boxes over which the bodies should be translated
to GDML involves first evaluating each region with very large \pyinline{Tub},
\pyinline{EllipticalTube} and \pyinline{Box} instances (by default
\SI{50}{\km} in length), such that they are effectively infinite for most
reasonable use cases.  \PYGEOMETRY{}'s CSG meshing is then used to generate
a mesh for each region from which the axis-aligned bounding box can be
extracted.  Each region is then evaluated a second time with respect to its
 bounding box, with all of its constituent infinite solids being
reduced in size as described above.  The implementation is described in
Algorithm~\ref{algo:redundant-halfspace}. This algorithm works robustly for
half-spaces, infinite circular cylinders and infinite elliptical cylinders
as the number of facets belonging to the mesh for such solids is
independent of the size of the solid, so generating the initial mesh from
arbitrarily large solids works well.  Generating a quadric surface over a very large
volume in space whilst retaining topological information is computationally
very expensive, so to resolve this the user must provide the approximate
axis-aligned bounding box of any region in which a quadric is used to
constrain the size of the mesh only to where it is needed.

@@

Any useful translation between geometry description formats must also
account for materials, and accordingly \PYGEOMETRY{} correctly translates
FLUKA materials to GDML.  FLUKA materials can be divided into built-in,
single-element and compound materials.  Built-in materials are simply those
that are predefined by FLUKA, single-element materials are described with a
single \fluka{material} card, and compound materials are described with one
\fluka{material} card followed by one or more \fluka{compound} cards.
These three alternatives are represented in \PYGEOMETRY{} with the
\pyinline{BuiltIn}, \pyinline{Material}, and \pyinline{Compound} classes.
Populating a hierarchy of instances using these classes is made more involved due to
the fact that recursively-defined materials in FLUKA input files need not
be defined in a logical order.  Namely, a given compound may be
defined before the materials that it consists of are themselves defined.
To account for this it is necessary to correctly compute the instantiation
order so that the above classes can be instantiated correctly.  To do this
a directed acyclic graph is populated with the materials and their
constituents, after which a topological sort is performed so that the
compound materials are sequenced after their constituents.  Mapping this
set of nested FLUKA materials to GDML materials is then
straight forwards as the two formats are similarly expressive.

@@

Figure~\ref{fig:fluka-to-geant4-cup} shows a Faraday cup implemented in
FLUKA and accurately translated to GDML using \PYGEOMETRY{}.  Many of the
steps described above were directly applied in this model and all the
features are tested and demonstrated in the repository.  This set of
algorithms for bridging FLUKA with GDML covers a very broad range of
geometries, however a number of possible improvements remain for the
future.
For example, FLUKA regions can in general be disconnected, that is
to say the resulting set of points defined by that region may consist of
two disconnected subvolumes.  In FLUKA this typically manifests itself in
the form of a disconnected unions, but it is also allowed for intersections
and subtractions to be disconnected.  In the case of translating
disconnected unions from FLUKA to GDML, this presents no difficulties as
the GDML multi-union solid is explicitly allowed to be formed from
disconnected parts.  However, it is forbidden for \pyinline{Intersection} or
\pyinline{Subtraction} solids to be disconnected in this way and is
therefore necessary to account for this difference between the two codes.
One solution to this problem would involve detecting and separating
disconnected meshes, and then placing each separated component as a
separate \pyinline{TesselatedSolid} instance.  This decomposition has not
been implemented and is left as future work.

@@

reader = freecad.Reader("CadFileName.step")
reader.relabelModel()
reader.convertFlat()
logical = reader.getRegistry().getWorldVolume()
\end{lstlisting}
Compared to other file readers, two additional steps are required: \pyinline{relabelModel} and \pyinline{convertFlat}. CAD model
part names can contain characters which are not allowed in Python
dictionaries so need to be replaced by using \pyinline{relabelModel}.
CAD models might also have a hierarchy of parts and assemblies, these are
converted without this structure by using \pyinline{convertFlat}.
In general there is no requirement to avoid geometric overlap  of parts in
a CAD file. This will result in overlaps between the converted tessellated
solids. This is avoided by shrinking each solid by computing the normal ${\bf n}$  for each
vertex ${\bf v}$ and shifting its position by $\epsilon {\bf n}$, so that
the new vertices are positioned at ${\bf v} - \epsilon {\bf n}$. The degree of shrinkage
is user-controllable.

>>>>

\linenumbers

@@

\begin{small}
\noindent
{\em Program Title: \PYGEOMETRY{} }                                         		\\
{\em Licensing provisions: GPLv3 }							\\
{\em Programming language: Python, C++}                         		\\
{\em External routines/libraries: ANTLR, CGAL, FreeCAD, NumPy, OpenCascade, SymPy, VTK} \\
%{\em Supplementary material:}                                 				\\
  % Fill in if necessary, otherwise leave out.
%{\em Journal reference of previous version:}                  			\\
  %Only required for a New Version summary, otherwise leave out.
%{\em Does the new version supersede the previous version?:}   	\\
  %Only required for a New Version summary, otherwise leave out.
%{\em Reasons for the new version:}							\\
  %Only required for a New Version summary, otherwise leave out.
%{\em Summary of revisions:}*								\\
  %Only required for a New Version summary, otherwise leave out.
{\em Nature of problem:}\\  % (approx. 50-250 words)
Creating computer-readable geometry descriptions for Monte Carlo radiation transport (MCRT) codes is a time-consuming and error-prone task.
Typically these geometries are written by the user directly in the file format used by the MCRT code. There are also multiple MCRT codes
available and geometry conversion is difficult or impossible between these simulation tools.
\\
{\em Solution method:}\\  % (approx. 50-250 words)
Create a Python application programming interface for the description and manipulation of Geant4 and FLUKA geometries, with full support for
the direct reading and writing of their respective geometry description file formats.
Form triangular meshes to represent geometric objects for both visualisation of the
geometry and to enable the use of advanced mesh-based geometric algorithms. Triangular mesh processing algorithms allow the loading and use of Standard Triangle Language (STL) and CAD/CAM
files. Converting from FLUKA to Geant4 requires algorithms to decompose solids to a set of unions of convex solids. Converting from
FLUKA to Geant4 requires a number of steps including the replacement of infinite surfaces with finite solids and the automatic elimination of overlaps.

@@

%% main text
\section{Introduction} \label{sec:introduction}
There are numerous software codes to simulate the passage of particles through material. Such radiation transport (RT) programs
include MCNP~\cite{Mcnp_Werner}, FLUKA~\cite{Fluka_Ferrari,Fluka_Bohlen}, Geant3~\cite{Geant3_Brun} and Geant4~\cite{Geant4_Agostinelli}.
All these codes are based on the Monte Carlo technique but each code either has a particular speciality, simulation methodology or target user community.
 Monte Carlo RT (MCRT) simulations have diverse uses including shielding calculations for radiological protection, detector performance, medical
imaging and therapy, and space radiation environment simulations. A fundamental requirement of all of the codes is to supply a
computer-readable description of the  physical 3D geometry that  the particles are passing through.  The creation of geometry description files is
typically a very time-consuming activity and the simulation validity and performance is directly dependent on the quality of the geometry. There is no
standard geometry format used across MCRT codes, with each code often using its own unique format. A user will typically not have geometry in both FLUKA and Geant4 for example. A geometry system
that enables the conversion between files prepared for different codes will allow for cross-checks of the physics processes in different particle transport
 codes.  The file formats used for geometry are generally focused
 on the computational efficiency of particle tracking algorithms  and not ease of preparation. In addition to the creation of geometry files for RT programs, usually
 computer-aided design (CAD) files exist for systems which need to be simulated. The fundamental geometric representations in CAD files are usually not
 amenable to MCRT programs.  For these reasons it is advantageous to create a software tool that allows particle transport code users to rapidly develop
 error-free geometry files, convert between common MCRT geometry formats and load CAD models.

This paper describes a geometry creation and conversion package called \PYGEOMETRY{}, written in Python and internally based on the Geant4 application
programming interface (API) and the Geometry Description Markup Language (GDML) for file persistency~\cite{GDML}. The main features of \PYGEOMETRY{}
are a Python scripting API to rapidly design parametrised geometry; conversion to and from  FLUKA geometry descriptions; conversion from CAD formats  (STEP
and IGES) based on FreeCAD~\cite{FreeCAD} and OpenCascade~\cite{OpenCASCADE}; and  geometry visualisation tools based on VTK~\cite{VTK4}. The
origin of \PYGEOMETRY{} was a set of utilities to prepare geometry for an accelerator beamline simulation program based on Geant4 called BDSIM~\cite{BDSIM_Nevay}.
Accelerator physicists, like specialists in other areas, need a tool to quickly model specialist geometry and the subsequent interaction of the charged particle beam.
\PYGEOMETRY{} allows the rapid creation and adaptation of geometry, with Figure~\ref{fig:workflow} demonstrating various possible workflows. \PYGEOMETRY{}
is not an executable software package but a toolkit. A user would typically write a very small Python program to use the classes and functions provided by
\PYGEOMETRY{}. This paper describes version 1.0 of \PYGEOMETRY{}, which is freely available as a Git repository and via the Python Package Index (PyPi).

@@

There are existing codes that have functionality similar to \PYGEOMETRY{}. ROOT~\cite{Brun:1997pa}, the high-energy physics data analysis framework,
can load, display and manipulate GDML-like geometry. ROOT is typically harder for the layperson (or non-high energy physicist) to use compared to the python implementation of \PYGEOMETRY{}. Existing open source tools to convert CAD files to GDML include GUIMesh~\cite{GUIMesh_Pinto}, with commercial
solutions including ESABASE2~\cite{ESABASE2} and FASTRAD~\cite{FASTRAD}. There are also tools from the fusion and neutronics community that can convert CAD
geometry into formats usable by MCRT codes, with examples including DAGMC~\cite{DAGMC} and McCAD~\cite{McCad}. In principle CAD software can export shape data to STL
(or other similar mesh formats), which can be used by Geant4~\cite{poole2012acad}. However, in practice using CAD models is difficult if that model is composed of a
large number of parts. This is because exporting, assigning material to, and placing the STL components into the MCRT code can be very cumbersome. Almost all modern CAD tools such as
CATIA, Inventor and SolidWorks have a scripting language to allow users to programmatically generate geometry. Similar CAD-style scripting languages do not exist for either GDML
or FLUKA and the existing set of software does not provide a complete set of tools to efficiently create complex geometries for these codes.

This paper is structured as follows: first comes a brief radiation transport-focused introduction to computer descriptions of geometry and an explanation of the design
and implementation of \PYGEOMETRY{}. Subsequent sections describe how \PYGEOMETRY{} can be used to perform rapid geometric modelling as well as conversions from
FLUKA to GDML, GDML to FLUKA and CAD to GDML. The paper concludes with an example of a composite, complex system consisting of components drawn from
all the supported  geometry input files formats.

\section{Computer descriptions of geometry} \label{sec:geometric}
Central to a computer-readable geometry is how a solid is defined in three dimensions. There are numerous ways to describe a
geometry, including constructive solid geometry (CSG), boundary representation (BREP) and tessellated polygons, which are described
briefly in this section.  The Geant4 geometry specification is a mixture of all three of these geometry modelling techniques and described in detail last.

%%%%%%%%%%%%%%%%% Constructive solid geometry (CSG)
Constructive solid geometry uses Boolean operations (subtraction, intersection and union) between simple solid shapes (e.g. cube, cylinder, sphere, etc.) or infinite
volumes (e.g. a plane-defined infinite half-space) to model complex surfaces, which represent a solid. Boolean operations and solids can be combined to form a
CSG tree to model complex geometry. FLUKA uses CSG to model solids. However the form of Boolean expression used by FLUKA is not a general CSG tree but a
logical expression in disjunctive normal form.

@@

\subsection{Visualisation} \label{sec:visualisation}
When implementing geometry, a rapid and robust visualisation system is key to produce error-free and efficient simulation input.
A  \PYGEOMETRY{} geometry hierarchy can be viewed using the popular Visualisation Toolkit (VTK). No separate scene graph is required as the Geant4
volume hierarchy is sufficient to place the meshes associated with each physical volume. A daughter volume is placed within a logical volume with a
rotation~$\mathbf{R}_d$, reflection~$\mathbf{S}_d$ and translation~$\mathbf{T}_d$.

@@

\subsection{Overlap detection}
All MCRT codes cannot handle spatial overlap between two geometric objects and will have ill-defined behaviour when tracking particles
in such a situation.  A key feature of \PYGEOMETRY{} is the detection of potential overlaps in a way which is most useful to the user. It does this by
performing an intersection operation between solid mesh instances and determining if the resulting mesh is empty. Figure~\ref{fig:overlap} shows three different types of possible
overlaps, (a) protrusion of a daughter from the mother, (b)  finite volume intersection between two daughters  and (c) an overlap where two daughters
share a face. If the resulting intersection is non-null then the overlaps can be displayed side-by-side in the visualisation.
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=8cm]{overlap}
\caption{Schematic of the three different types of overlap between the daughters of a mother logical volume.}
\label{fig:overlap}
\end{center}
\end{figure}

@@

Overlap detection in \PYGEOMETRY{} relies on the meshes generated for each solid.  As
these mesh representations will generally approximate their respective solid,
so to will any overlap detection algorithm be an approximation. Generally, the overlap detection
algorithm proceeds as shown in Algorithm~\ref{algo:overlap}.
For a logical volume with $n_{\rm daughters}$ physical volumes, assuming meshes for solids have an average number of faces $n$, clearly this algorithm
has complexity $\sim \mathcal{O}(n^2 n_{\rm daughters}^2)$. This complexity
can result in a steep computational cost for highly-flat geometry
hierarchies, but overall remains worthwhile considering the potential
waste in large amounts of cluster CPU time if small overlaps are present in the final MCRT simulation. This algorithm clearly favours geometry
descriptions which have a high degree of logical volume reuse, which this is also true of Geant4 as a whole.
Due to the discrete nature of triangular meshes it is not possible to have perfect detection
of overlaps, especially when curved surfaces are considered. In some rare
cases either overlaps maybe missed, or spurious overlaps maybe reported.  However, the density of the meshes
created for the solids is controllable by the user, so the user can opt for more precise overlap checking at the cost of greater computation time.  
In general, the overlap detection algorithm can present the potential overlaps quickly and easily to the user, so significantly aid the geometry 
debugging process.

@@

# physical volume visualization attributes
bp1.visOptions.color = (1,0,0)
bp1.visOptions.alpha = 1.0
bp2.visOptions.color = (0,1,0)
bp2.visOptions.alpha = 1.0
bp3.visOptions.color = (0,0,1)
bp3.visOptions.alpha = 1.0

@@

The \PYGEOMETRY{} Python code in the example is approximately as expressive
as the GDML it writes. The benefit of wrapping GDML in Python is that it allows
 very rapid prototyping of geometry without the overhead of C++ compilation  (in the
case of implementing the geometry directly in Geant4) or writing well-formed XML (in the case of GDML). Effectively,
by using \PYGEOMETRY{}, the set of possible user-errors when describing geometry in C++ or XML
instead either manifest as Python exceptions or are eliminated entirely.
Another key benefit is the ability to use the \PYGEOMETRY{} code to create
programmatic converters between different geometry languages or more generally
manipulation and transformation of the geometry stored in memory. The rapid modelling
example given in Listing~\ref{lst:pythonRapidModelling} and Figure~\ref{fig:rapidModellingExample}
is rather trivial. A significantly more complex example is shown in Figure~\ref{fig:gdml-flair}.

@@

Generating these bounding boxes over which the bodies should be translated
to GDML involves first evaluating each region with very large \pyinline{Tub},
\pyinline{EllipticalTube} and \pyinline{Box} instances (user adjustable but default
\SI{50}{\km} in length, ), such that they are effectively infinite for most
reasonable use cases.  \PYGEOMETRY{}'s CSG meshing is then used to generate
a mesh for each region from which the axis-aligned bounding box can be
extracted.  Each region is then evaluated a second time with respect to its
 bounding box, with all of its constituent infinite solids being
reduced in size as described above.  The implementation is described in
Algorithm~\ref{algo:redundant-halfspace}. This algorithm works robustly for
half-spaces, infinite circular cylinders and infinite elliptical cylinders
as the number of facets belonging to the mesh for such solids is
independent of the size of the solid, so generating the initial mesh from
arbitrarily large solids works well.  Generating a quadric surface over a very large
volume in space whilst retaining topological information is computationally
very expensive, so to resolve this the user must provide the approximate
axis-aligned bounding box of any region in which a quadric is used to
constrain the size of the mesh only to where it is needed.

@@

Any useful translation between geometry description formats must also
account for materials, and accordingly \PYGEOMETRY{} correctly translates
FLUKA materials to GDML.  FLUKA materials can be divided into built-in,
single-element and compound materials.  Built-in materials are simply those
that are predefined by FLUKA, single-element materials are described with a
single \fluka{material} card, and compound materials are described with one
\fluka{material} card followed by one or more \fluka{compound} cards.
These three alternatives are represented in \PYGEOMETRY{} with the
\pyinline{BuiltIn}, \pyinline{Material}, and \pyinline{Compound} classes.
Populating a hierarchy of instances using these classes is made more involved due to
the fact that recursively-defined materials in FLUKA input files need not
be defined in a logical order.  Namely, a given compound may be
defined before the materials that it consists of are themselves defined.
To account for this it is necessary to correctly compute the instantiation
order so that the above classes can be instantiated correctly.  To do this
a directed acyclic graph is populated with the materials and their
constituents, after which a topological sort is performed so that the
compound materials are sequenced after their constituents.  Mapping this
set of nested FLUKA materials to GDML materials is then
straightforward as the two formats are similarly expressive.

@@

Figure~\ref{fig:fluka-to-geant4-cup} shows a Faraday cup implemented in
FLUKA and accurately translated to GDML using \PYGEOMETRY{}.  Many of the
steps described above were directly applied in this model and all the
features are tested and demonstrated in the repository.  This set of
algorithms for bridging FLUKA with GDML covers a very broad range of
geometries. However a number of possible improvements remain for the
future.
For example, FLUKA regions can in general be disconnected, that is
to say the resulting set of points defined by that region may consist of
two disconnected subvolumes.  In FLUKA this typically manifests itself in
the form of a disconnected unions, but it is also allowed for intersections
and subtractions to be disconnected.  In the case of translating
disconnected unions from FLUKA to GDML, this presents no difficulties as
the GDML multi-union solid is explicitly allowed to be formed from
disconnected parts.  However, it is forbidden for \pyinline{Intersection} or
\pyinline{Subtraction} solids to be disconnected in this way and is
therefore necessary to account for this difference between the two codes.
One solution to this problem would involve detecting and separating
disconnected meshes, and then placing each separated component as a
separate \pyinline{TesselatedSolid} instance.  This decomposition has not
been implemented and is left as future work.

@@

reader = freecad.Reader("CadFileName.step")
reader.relabelModel()
reader.convertFlat()
logical = reader.getRegistry().getWorldVolume()
\end{lstlisting}
Compared to other file readers, two additional steps are required: \pyinline{relabelModel} and \pyinline{convertFlat}. CAD model
part names can contain characters which are not allowed in Python
dictionaries so need to be replaced by using \pyinline{relabelModel}.
CAD models might also have a hierarchy of parts and assemblies. These are
converted without this structure by using \pyinline{convertFlat}.
In general there is no requirement to avoid geometric overlap  of parts in
a CAD file. This will result in overlaps between the converted tessellated
solids. This is avoided by shrinking each solid by computing the normal ${\bf n}$  for each
vertex ${\bf v}$ and shifting its position by $\epsilon {\bf n}$, so that
the new vertices are positioned at ${\bf v} - \epsilon {\bf n}$. The degree of shrinkage
is user-controllable.

An example geometry representing a dipole electromagnet consisting of three parts was created in a commercial CAD  package (Autodesk Fusion 360)
and saved as a STEP file. The \PYGEOMETRY{}-produced output is shown in Figure~\ref{fig:cad-gdml}.
